---
title: "Low Birth Rates Analysis_US"
output: html_notebook
---
#Introduction 

What is the relationship between low birth rates and smoking and alcohol consumption in the U.S.? It is important to note that is aggregate data and not individual cases.
```{r}
library(nlme)
library(dplyr)
library(magrittr)
library(tidyr)
library(haven)
library(lmerTest)
library(ggplot2)
library(texreg)
```

```{r}
library(radiant.data)
library(readr)
  lbr1 <- read_csv("/Users/racruz/Desktop/lbr.csv", col_names = TRUE)
   head(lbr1)
```

#Preparing Data for Analysis

```{r}

lbr1 <- rename(lbr1, state = Geo_STATE, county = Geo_COUNTY, lowbirth = SE_NV002_001,smokers = SE_T011_001,alcohol = SE_T011_002)

```

#Data Cleaning
Originally the data contained 3141 observations. After running na.omit the leftover cases were 3041 observations.
```{r}
lbr1<-na.omit(lbr1)
head(lbr1)
```

#Preliminary Investigation of "Low Birth Rate In The U.S"

```{r}
ggplot(lbr1, aes(x = lowbirth, y = state)) + geom_jitter(alpha = 0.5)

ggplot(lbr1, aes(x = lowbirth, y = smokers*alcohol)) + geom_jitter(alpha = 0.5)

ggplot(lbr1, aes(x = lowbirth, y = smokers, color = "Respondents")) + geom_jitter(alpha = 0.5)

ggplot(lbr1, aes(x = lowbirth, y = alcohol, color = "Respondents")) + geom_jitter(alpha = 0.5)

```

#MULTI-LEVEL ANALYSIS X RANDOM EFFECTS
In understanding low birth rates in the U.S I am using state as higher level observation.


#Complete pooling
Model 1 displays that the lowbirth intercept is 2416.17. Smokers have less of an impact on low birth rate than alcohol consumers with a value of (-95.11). In contrast alcohol consumers have more of an impact on low birth rates in the U.S with a value of 6.19.

```{r}
 lbreg <- lm(lowbirth ~ smokers + alcohol, data = lbr1)
screenreg(lbreg)
visreg::visreg(lbreg)
```

```{r, message=FALSE, warning=FALSE, results="asis"}
htmlreg(list(lbreg))
```


#Smokers X Alcohol Interaction
As we see this model displays a low birth intercept of 1497.37. Smokers have less impact on low birth weight than alcohol consumers with a value of (-41.50) and alcohol consumers with a value of (70.46). When the interacton term is introduced we see that smokers and alcohol cosumers have a slightly negative impact on low birth weights in the U.S.

```{r}
lbr3 <- gls(lowbirth ~ smokers*alcohol, data = lbr1, method = "ML")
screenreg(lbr3)
plot(lbr3)

```


```{r, message=FALSE, warning=FALSE, results="asis"}
htmlreg(list(lbr3))
```


#No Pooling
Intercept of low birth rates in the U.S range from 1,000 to 5,000 by states when looking at smokers.

```{r}

dcoef <- lbr1 %>% 
    group_by(state) %>% 
    do(mod = lm(lowbirth ~ smokers, data = .))
coef <- dcoef %>% do(data.frame(intc = coef(.$mod)[1]))
ggplot(coef, aes(x = intc)) + geom_histogram(color = "black", fill = "aqua marine1")

```


#No plotting: Plotting the slopes
The relationship between smokers and low birth weight  will vary by state. Some will display a more a negative relationship and others will display a negative relationship however, it seems in this distribution it will more likely display a negative intercept.


```{r}
dcoef <- lbr1 %>% 
    group_by(state) %>% 
    do(mod = lm(lowbirth ~ smokers, data = .))
coef <- dcoef %>% do(data.frame(smokers = coef(.$mod)[2]))
ggplot(coef, aes(x = smokers)) + geom_histogram(color = "orange", fill = "purple")
```






#Partial Pooling
Accounting for lowbirth weight variations between states. We see consistent relationships with the models above. 


```{r}
m1_lme <- lme(lowbirth ~ smokers, data = lbr1, random = ~1|state, method = "ML")
summary(m1_lme)
screenreg(m1_lme)
```


```{r}
m2_lme <- lme(lowbirth ~ alcohol, data = lbr1, random = ~ alcohol|state, method = "ML")
summary(m2_lme)
screenreg(m2_lme)

```

```{r}
AIC(lbr3, m1_lme, m2_lme)
```

```{r}
m3_lme <- lme(lowbirth ~ smokers + alcohol, data = lbr1, random = ~ alcohol|state, method = "ML")
summary(m3_lme)
screenreg(m3_lme)
```

```{r}
m4_lme <- lme(lowbirth ~ smokers*alcohol, data = lbr1, random = ~ alcohol|state, method = "ML")
summary(m4_lme)
plot(m4_lme)
screenreg(m4_lme)
```
#Model Fit
Strongest model that fits this data seems to be m3_lme.
```{r}
AIC(lbr3, m1_lme, m2_lme, m3_lme, m4_lme)

```


